---
title: "Dashboard of Interactive Time Series Plots"
author: "Jerzy Pawlowski at NYU Tandon"
affiliation: NYU Tandon School of Engineering
email: jp3900@nyu.edu
date: '`r format(Sys.time(), "%m-%d-%Y %H:%M")`'
output:
  html_document:
    fig_width: 6
    fig_height: 4
---

```{r setup, include=FALSE}
# This is an R setup chunk, containing default options applied to all other chunks
knitr::opts_chunk$set(echo=FALSE)
library(rutils)
library(dygraphs)
```


### A dashboard of interactive time series plots written in *Rmarkdown*  

This *Rmarkdown* document contains interactive *dygraphs* plots of time series contained in the package *rutils*.  The interactive time series plots are produced using the package *dygraphs*.  It creates a list of *dygraphs* objects in a loop, and then uses the package *htmltools* to render (plot) the *dygraphs* objects.  

### You can click and drag your mouse to zoom into any plot.  

Double-click your mouse to restore the plot.


```{r echo=FALSE}
# extract the closing prices into a list of xts time series
price_s <- lapply(rutils::env_etf$sym_bols, function(sym_bol) {
  quantmod::Cl(get(sym_bol, envir=rutils::env_etf))
})  # end lapply

# flatten (cbind) list of prices into single xts series
price_s <- rutils::do_call(cbind, price_s)

# overwrite NA values
price_s <- rutils::na_locf(price_s)
price_s <- rutils::na_locf(price_s, from_last=TRUE)

# make nice column names
colnames(price_s) <- rutils::do_call_rbind(strsplit(colnames(price_s), split="[.]"))[, 1]

# create a list of dygraphs objects in loop
dy_graphs <- lapply(price_s, function(pric_e) {
  dygraphs::dygraph(pric_e, main=paste("Plot of:", colnames(pric_e)), width=600, height=400)
})  # end lapply

# render the dygraphs objects
htmltools::tagList(dy_graphs)
```
