---
title: "Kaggle Credit Card Defaults in Taiwan in 2005"
author: "Jerzy Pawlowski"
affiliation: NYU Tandon School of Engineering
email: jp3900@nyu.edu
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output:
  html_document:
    code_folding: hide
    theme: journal
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Default of Credit Card Clients Dataset {.tabset}

### Overview

<br> 

**Dataset Information**

This dataset contains information on default payments, demographic factors, credit data, history of payment, and bill statements of credit card clients in Taiwan from April 2005 to September 2005.

```{r setup, echo=FALSE, include=FALSE}
library(ggplot2)
library(dplyr)
library(corrplot)
library(gridExtra)
library(caret)
library(e1071)
library(gridBase) # Draw Viz
library(gtable) # Viz
library(vcd) # mosaic
library(Hmisc) # rcorr
library(Amelia) # missmap
library(ROCR) 
library(rpart)
library(randomForest)


options("scipen" = 10) # make sure all the plot have no exponential mark -> show real values


raw.credit <- read.csv("../input/UCI_Credit_Card.csv", header=T, sep=',', stringsAsFactors=F)
colnames(raw.credit) <- tolower(colnames(raw.credit))
```

I have put the column names to lower cases. Let's quickly look at the dataset.

* There're quite lots of features in dataset (25 columns)
* **id** column is the client identifier, this is not important and will be removed.
* There're few variables which I will convert to factor (such as sex, education, marriage). pay_0 and other pay_x features can considered to be factor variable, but let's keep them as numeric for now.

```{r sample}
dim(raw.credit)
summary(raw.credit)
head(raw.credit,1)
```

Is there any missing data points in the dataset?

* There are no missing data.

```{r check missing, warning=FALSE}
missmap(raw.credit, main = "Deafult credit card missing data heatmap", 
        col=c("red", "black"), legend=F)

colSums(sapply(raw.credit, is.na))
```

Note that we have removed **id** from datasets as there's no meaning in this context.
We will proceed to next section before assigning any factor class to the variables.

```{r convert data, echo=FALSE}
convert.to.factor <- function(df, lst){
  for (i in lst){
    df[[i]] <- as.factor(df[[i]])
  }
  return(df)
}

convert.to.numeric <- function(df, lst){
  for (i in lst){
    df[[i]] <- as.numeric(df[[i]])
  }
  return(df)
}


raw.credit <- select(raw.credit,
                     -id)
```

### Basic Visual & Data Cleansing

This section will try to make more senses on each data point by visualization. I will also fix the data points if they are not making any senses from the given definition.

```{r Utility, echo=FALSE}
barPlot <- function(df, i) {
  d <- data.frame(x=df[[i]]) 
  p <- ggplot(d, aes(factor(x))) + 
    stat_count() + 
    xlab(colnames(df)[i]) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)
          , legend.position = "bottom"
          , legend.title = element_blank()
    )  
  return(p)
}

histPlot <- function(df, i) {
  d <- data.frame(x=df[[i]])
  p <- ggplot(d, aes(x)) +
    geom_histogram(alpha=0.5) +
    xlab(colnames(df)[i]) +
    theme_bw() +
    theme(legend.position = "right",
          legend.title = element_blank()
          )
  return(p)
}


plotDen <- function(df, i){
  data <- data.frame(x=df[[i]], limit_bal = df$limit_bal)
  medx <- median(df[[i]])
  p <- ggplot(data= data) + 
    geom_line(aes(x = x), stat = 'density', size = 1,alpha = 1.0) +
    xlab(paste0((colnames(df)[i]), '\n', 'Median: ',round(median(df[[i]], na.rm = TRUE), 2))) +
    geom_vline(xintercept = medx, linetype="dashed", size=0.5) +
    theme_bw() 
  return(p)
}

lets.plot <- function(df, fun, ii, ncol) {
  pp <- list()
  for (i in ii) {
    p <- fun(df=df, i=i)
    pp <- c(pp, list(p))
  }
  do.call("grid.arrange", c(pp, ncol=ncol))
}
```

1. Look at Demographical variable data
+ There are more female than male in the dataset.
+ There are clients who finished university-level education.
+ There are more single client than married, but the number is quite closed (53.2%:45.5%)

Recap from data definition:

* EDUCATION: (1=graduate school, 2=university, 3=high school, 4=others, 5=unknown, 6=unknown)
* MARRIAGE: Marital status (1=married, 2=single, 3=others)

```{r Viz1}
# Plot Demographical data
lets.plot(raw.credit, fun=barPlot, ii=2:4, ncol=2)

```

Let's fix the datapoint of education and marriage.

* For education, we will merge 0, 4, 5, 6 to 4 (others)
* For marriage, we will merge 0 to 3 (others)

```{r data cleansing 1, echo=FALSE}
raw.credit[which(raw.credit$education %in% c(0, 4, 5, 6)),]$education <- 4 # Flag to others
raw.credit[which(raw.credit$marriage == 0),]$marriage <- 3 # Flag to other

list.to.factor <- c("sex", "education", "marriage", "default.payment.next.month")
raw.credit <- convert.to.factor(raw.credit, list.to.factor)

lets.plot(raw.credit, fun=barPlot, ii=3:4, ncol=2)
```

2. Age data - the data looks ok, 

```{r age viz}
lets.plot(raw.credit, fun=plotDen, ii=5, ncol=1)
```

3. Quick look on the limit balance

```{r Viz2}
lets.plot(raw.credit, fun=plotDen, ii=1, ncol=1)

```

4. What about repayment status

Based on the data definition, we have the following:

* -1 = pay duly
* 1 = payment delay for one month
* 2 = payment delay for two months 

.
..

* 8 = payment delay for eight months

and so on ... 

However, from the plot, we can see that there are a lot of -2 and 0 data points spreading out across repayment status.

With this, I **assume** that -2, -1 and 0 have the same meaning. We will merge them to 0.

```{r repayment}
lets.plot(raw.credit, fun=barPlot, ii=6:11, ncol=2)

for (i in seq(6, 11, 1)) {
  raw.credit[which(raw.credit[[i]] %in% c(-2, -1, 0)),][[i]] <- 0
}
```

5. Let's quickly look at the bill statement and previous payment in each month.

```{r bill pay plot}
lets.plot(raw.credit, fun=plotDen, ii=12:17, ncol=2)

lets.plot(raw.credit, fun=plotDen, ii=18:23, ncol=2)

```

6. Lastly, let's look at the number of default payment next month.

The default payment is around 22% of the total observations in the dataset.

```{r default plot}
lets.plot(raw.credit, fun=barPlot, ii=24, ncol=1)
prop.table(table(raw.credit$default.payment.next.month))
```

Let's move to the next section.

### Correlation

Let's explore the correlation between each data points.

1. I will use the mosaicplot to visualize the demographical data against the default payment.

```{r gender edu martial}

mosaicplot(raw.credit$sex ~ raw.credit$default.payment.next.month, shade=F, color=T,
           main="Gender affecting the default payment next month?",
           ylab="Default payment next month (1-Y, 0-N)",
           xlab="Gender (1-male, 2-female)"
          )

mosaicplot(raw.credit$education ~ raw.credit$default.payment.next.month, shade=F, color=T,
           main="Education affecting the default payment next month?",
           ylab="Default payment next month (1-Y, 0-N)",
           xlab="Education  (1=graduate school, 2=university, 3=high school, 4=others)"
          )

mosaicplot(raw.credit$marriage ~ raw.credit$default.payment.next.month, shade=F, color=T,
           main="Marital status affecting the default payment next month?",
           ylab="Default payment next month (1-Y, 0-N)",
           xlab="Marriage (1=married, 2=single, 3=others)"
          )

mosaicplot(paste0(raw.credit$sex, '-', raw.credit$marriage) ~ raw.credit$default.payment.next.month,
           shade =F, color = T, main= "Gender with marital status against default credit?",
           xlab="Gender - marital status", ylab="Default payment next month (1-Y, 0-N)")

```

2. Let's combine everything

It looks like the married women are more likely to default payment.


```{r combine}
ggplot(raw.credit, aes(education, fill=default.payment.next.month)) +
  stat_count() +
  facet_wrap(~paste0(raw.credit$sex, '-', raw.credit$marriage)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)
          , legend.position = "bottom"
          , legend.title = element_blank()
        )  
```


3. Look at the repayment status of each month

From pay_0 to pay_6, the definition is the same except it is reflected the repayment status of each month (-1=pay duly, 1=payment delay for one month, 2=payment delay for two months, ... 8=payment delay for eight months, 9=payment delay for nine months and above).

```{r repayment-to-default}
repay.bar.plot <- function(df, i, target) {
  d <- data.frame(x=df[[i]]) 
  p <- ggplot(d, aes(factor(x), fill=df[[target]])) + 
    stat_count() + 
    xlab(colnames(df)[i]) + 
    theme_bw() +
    theme(axis.text.x = element_text(angle = 0, hjust = 1)
          , legend.position = "right"
          , legend.title = element_blank()
    )  
  return(p)
}


repayment.plot <- function(df, fun, ii, target, ncol) {
  pp <- list()
  for (i in ii) {
    p <- fun(df=df, i=i, target=target)
    pp <- c(pp, list(p))
  }
  do.call("grid.arrange", c(pp, ncol=ncol))
}

repayment.plot(raw.credit, fun=repay.bar.plot, ii=6:11, target=24, ncol=2)

```

Let's plot some corrplot to see if the data has any correlation.

```{r prep2, echo=FALSE, warning=FALSE, message=FALSE}
library(corrplot)
library(Hmisc) # for finding correlation and p-value

tmp <- read.csv("../input/UCI_Credit_Card.csv", header=T, sep=',', stringsAsFactors=F)
colnames(tmp) <- tolower(colnames(tmp))
tmp <- select(tmp,
              -id)

```


The figure and color are related to correlation coefficients with the number is the p-values.
Any sig.level less than or equal to 0.01, it will be crossed out.

```{r corrplot}
credit.corr <- rcorr(as.matrix(tmp), type=c("pearson"))
colnames(credit.corr$r) <- c(1:length(colnames(credit.corr$r)))
colnames(credit.corr$P) <- c(1:length(colnames(credit.corr$P)))

corrplot(credit.corr$r, order="FPC", type="upper", method="square",
         tl.col="black", tl.srt = 45, p.mat = credit.corr$P, addCoef.col = "black",
         number.cex = 0.5, sig.level = 0.01)
```

```{r remove some obj, echo=FALSE, message=FALSE}
rm(credit.corr, i, list.to.factor, tmp)
```

Not much correlation to our binary **'default.payment.next.month'** output. Let's dive to the modeling part.

### Model Building

This section is to start building the model. Several model will be attempted to build and predict the default payment output.

Separate the dataset onto training (70%) and testing (30%).
```{r prep train test, eval=FALSE}
intraining <- createDataPartition(y=raw.credit$default.payment.next.month, p=0.7, list=F)
train.batch <- raw.credit[intraining,]
test.batch <- raw.credit[-intraining,]
```

```{r outp dist, echo=FALSE}
set.seed(5012)
intraining <- createDataPartition(y=raw.credit$default.payment.next.month, p=0.7, list=F)
train.batch <- raw.credit[intraining,]
test.batch <- raw.credit[-intraining,]

cat("--- Distribution of training dataset --")
table(train.batch$default.payment.next.month)
round(prop.table(table(train.batch$default.payment.next.month)), 2)

cat("\n--- Distribution of training dataset --")
table(test.batch$default.payment.next.month)
round(prop.table(table(test.batch$default.payment.next.month)), 2)
```

1. Traditional Score using Logistic Regression

```{r glm, warning=FALSE}

glm.m <- glm(default.payment.next.month ~ ., data=train.batch, family=binomial)
summary(glm.m)

glm.final <- glm(default.payment.next.month ~ limit_bal+sex+education+marriage+
                   pay_0+pay_2+pay_3+pay_5+pay_6+pay_amt1+pay_amt2, data=train.batch,
                 family=binomial)
summary(glm.final)

```

```{r roc glm}
test.x <- select(test.batch, - default.payment.next.month)
test.batch$glm_scor1 <- predict(glm.final, type="response", test.x)
pred <- prediction(test.batch$glm_scor1, test.batch$default.payment.next.month)
glm.perf <- performance(pred, "tpr", "fpr")

ggplot(test.batch, aes(default.payment.next.month, glm_scor1)) +
  geom_boxplot() + theme_bw() +
  ggtitle("GLM Score distribution") +
  xlab("Default Payment") +
  ylab("Default Score")

# ROC
plot(glm.perf, lty=1, col="blue", main="Logistic Regression Performance")
abline(a=0, b= 1)

auc.perf <- performance(pred, measure = "auc")
acc.perf <- performance(pred, measure = "acc")
# AUC
cat(paste0("AUC:: ", auc.perf@y.values), "\n")

# Get accuracy and cutoff
ind <- which.max( slot(acc.perf, "y.values")[[1]] )
acc <- slot(acc.perf, "y.values")[[1]][ind]
cutoff <- slot(acc.perf, "x.values")[[1]][ind]

print(c(accuracy= acc, cutoff = cutoff))

# Draw Accuracy with cutoff
plot(acc.perf)
abline(v=cutoff)

glm.scor <- select(test.batch,
                   glm_scor1, 
                   default.payment.next.month)

glm.scor <- mutate(glm.scor,
                   predict_class = ifelse(glm_scor1 >= cutoff, 1, 0))

confusionMatrix(glm.scor$predict_class, glm.scor$default.payment.next.month, positive = '1')

```

Eventually, we can see which top X variables are influenced the predicted score. In this example, I will select only top 2 variables.

```{r glm terms}
glm.terms <-predict(glm.final,type='terms',test.x)

get.top.var <- function(x,top=2){ 
  res=names(x)[order(x, decreasing = TRUE)][1:top]
  paste(res,collapse=";",sep="")
}

top.var <- apply(glm.terms, 1, get.top.var, top=2)
test.batch %>% select(default.payment.next.month, glm_scor1) %>%
  cbind(top.var) %>% head(5)

```

```{r clean space, echo=FALSE, include=FALSE}
rm(top.var, glm.terms)
```



2. Decision Tree **(rpart library)**


```{r rpart, warning=FALSE}
rpart.model <- rpart(default.payment.next.month ~.,data=train.batch)
              
plot(rpart.model);text(rpart.model)
plotcp(rpart.model)
printcp(rpart.model)

rpart.pred <- predict(rpart.model, test.x, type="class")
confusionMatrix(rpart.pred, test.batch$default.payment.next.month, positive = '1')

test.batch$rpart_scor1 <- predict(rpart.model,type='prob',test.x)
pred2 <- prediction(test.batch$rpart_scor1[, 2], test.batch$default.payment.next.month)
rpart.perf1 <- performance(pred2,"tpr","fpr")


plot(rpart.perf1, col='red', main='Rpart Tree performance')
abline(a=0, b= 1)

```

3. Random Forest

```{r randomForest, warning=FALSE, message=FALSE}

rand.model <- randomForest(default.payment.next.month~.,data=train.batch, importance=T,
                            ntree=500, keep.forest=T)

varImpPlot(rand.model)

rand.pred <- predict(rand.model, test.x, type="class")
confusionMatrix(rand.pred, test.batch$default.payment.next.month, positive='1')

rand.prob <- predict(rand.model, test.x, type='prob')[,2]
pred4 <- prediction(rand.prob,test.batch$default.payment.next.month)

rf.perf1 <- performance(pred4,"tpr","fpr")

plot(glm.perf,col='red',lty=1, main='ROC Logistic VS. Rpart tree VS. Random Forest'); 
plot(rpart.perf1, col='blue',lty=2,add=T); 
plot(rf.perf1, col='green', lty=3, add=T);
legend(0.6,0.6,c('logistic regression','rpart tree', 'randomForest'),
       col=c('red','blue','green'),lwd=3)


```

Final print the AUC of all 3 predictions...
```{r AUC}
#simple model
glm.auc <- performance(pred,"auc")
cat("\n=== Logistic Regression ===\n")
cat(paste0(glm.auc@y.name,":: ", round(glm.auc@y.values[[1]], 2)))
cat(paste0("\nKS :: ", 
           round(max(attr(glm.perf,'y.values')[[1]]-attr(glm.perf,'x.values')[[1]]),2)))

# rpart
rpart.auc <- performance(pred2,"auc")
cat("\n\n=== Rpart tree ===\n")
cat(paste0(rpart.auc@y.name,":: ", round(rpart.auc@y.values[[1]],2)))
cat(paste0("\nKS :: ", 
           round(max(attr(rpart.perf1,'y.values')[[1]]-attr(rpart.perf1,'x.values')[[1]]),2)))

# RF
rf.auc <- performance(pred4,"auc")
cat("\n\n=== Random Forest ===\n")
cat(paste0(rf.auc@y.name,":: ", round(rf.auc@y.values[[1]], 2)))
cat(paste0("\nKS :: ", 
           round(max(attr(rf.perf1,'y.values')[[1]]-attr(rf.perf1,'x.values')[[1]]),2)))


```

